---
title: "Cnemaspis Analysis"
author: "Arka Pal"
date: "2023-08-01"
output:
  html_notebook: default
  html_document:
    df_print: paged
---

This notebook includes all statistical analyses and figures corresponding to the manuscript.

```{r setup, include=F}
knitr::opts_chunk$set(cache = T, echo = F)
knitr::opts_knit$set(root.dir = "~/Projects/2015-Cnemaspis/")
```

```{r Load libraries, include=F}
#rm(list=ls())
#libs <- c("tidyverse", "gridExtra", "knitr", "glmmADMB", "multcomp")
#install.packages(setdiff(libs,installed.packages()))
#lapply(libs, require, character.only=T)

#install.packages("glmmADMB", repos = "http://R-Forge.R-project.org") ### WRONG repo link
#install.packages("glmmADMB", repos = "http://glmmadmb.r-forge.r-project.org/repos") ### CORRECT repo link

library(tidyverse)
library(gridExtra)
library(glmmADMB)
library(multcomp)
library(lme4)
library(car) #for significance testing (anova)
library(performance)

my_geom_theme <- theme_bw() + theme(panel.grid = element_blank())
```

## 1. Variation in ectoparasite levels
```{r load-morphoData}
morph <- read.csv("./analysis/datasets/MorphologyDataNew.csv", header = T, stringsAsFactors = T)
morph$Month <- as.factor(morph$Month)
str(morph) #exploring data
```

```{r subset high-density}
## Subset lizards from only high-density plots
morph_highDen <- morph %>% filter(Plot_type == "H" & LizardType != "W")
## subset only specific morphs
morph_highDen_Ymales <- morph_highDen %>% filter(LizardType == "Y")
morph_highDen_females <- morph_highDen %>% filter(LizardType == "F")
```

```{r figS1, fig.width=4, fig.height=4, fig.align='center', echo=T, results='markup'}
#### Distribution of parasite count
#figs1
hist(morph_highDen$Parasites, xlab="Ectoparasite Level", ylab="Frequency", main="Distribution of parasite count", breaks = 40)
#data is zero biased, so using glmmadmb package

#### Test for normality
shapiro.test(morph_highDen$Parasites)
```
Fig S1: Histogram of ectoparasite level
<br/>

## 2. Measuring body condition 

#### Calculate body condition as scaled mass index (SMI)
```{r SMI_Ymales, echo=T, results='markup'}
## Males
# Population mean SVL
meanSVL_yM <- mean(morph_highDen_Ymales$SVL)
# Linear regression of log transformed mass by SVL
lm_yM <- lm(log(SVL)~log(Mass), morph_highDen_Ymales)
summary(lm_yM)
```

```{r SMI_Ymales_data}
slope_yM <- 0.05094
morph_highDen_Ymales$SMI <- morph_highDen_Ymales$Mass*(meanSVL_yM/morph_highDen_Ymales$SVL)^slope_yM
```

```{r SMI_females, echo=T, results='markup'}
## Females
# Population mean SVL
meanSVL_F <- mean(morph_highDen_females$SVL)
# Linear regression of log transformed mass by SVL
lm_F <- lm(log(SVL)~log(Mass), morph_highDen_females)
summary(lm_F)
```
```{r SMI_females_data}
slope_F <- 0.11503
morph_highDen_females$SMI <- morph_highDen_females$Mass*(meanSVL_F/morph_highDen_females$SVL)^slope_F
```

```{r SMI_data}
## Merge the male and female files 
morph_highDen <- rbind(morph_highDen_Ymales, morph_highDen_females)
str(morph_highDen)
```

## 3. Relationship between ectoparasites and body condition

```{r fig1, fig.height=4, fig.width=4, echo=F}
fig1 <- ggplot (morph_highDen, aes(x=Parasites, y=SMI, fill=LizardType)) +
  geom_point(shape=1, size=1, alpha=1, aes(colour=LizardType)) +
  geom_smooth(method='glm', fullrange = TRUE, se=TRUE, level=0.95, span=1, aes(colour=LizardType)) +
  labs(y="Body Condition (Scaled Mass Index)", x="Ectoparasite Level") + my_geom_theme +
  theme(legend.position = c(0.8,0.8), legend.direction = "vertical", legend.title = element_blank()) + 
  #ylim(-2,25) +
  scale_color_manual(labels=c("Females","Males"), values = c("grey30","gold3")) +
  scale_fill_manual(labels=c("Females","Males"), values = c("grey30","gold3"))

fig1
```
Fig1. Ectoparasite Level vs Body Condition 

<br/>
```{r figS2, fig.height=3, fig.width=6, echo=F}
## fig S2a
figS2a <- ggplot(morph_highDen, aes(x=LizardType, y=Parasites, fill=LizardType)) +
  geom_boxplot(size=1) +
  scale_x_discrete(labels=c("Female","Male")) +
  labs(x="Sex",y="Ectoparasite Level") +
  my_geom_theme + theme(legend.position="none") + ylim(-1,25) +
  scale_fill_manual(values = c("grey60","gold3"))


## fig S2b
months <- c("Sep","Oct","Nov","Dec","Jan","Feb","Mar")

figS2b <- ggplot(morph_highDen, aes(x=Month, y=Parasites)) + 
  geom_boxplot(size=1) +
  scale_x_discrete(labels=c("Sep","Oct", "Nov", "Dec", "Jan", "Feb", "Mar")) +
  labs(x="Months", y="") +
  my_geom_theme +
  ylim(-1,25)

grid.arrange(figS2a, figS2b, widths=c(0.35,1), nrow=1)
```
FigS2. Ectoparasite level across sex and months. 


<br/>
<br/>

#### Table S1
```{r TableS1: Model for Body Condition, warning=F, echo=TRUE, results='markup'}
ParVarMod <-  glmmadmb(Parasites ~ LizardType * SMI + Month,
                       random = ~1|PlotID , 
                       family = "poisson", 
                       zeroInflation = T,
                       data = morph_highDen)
summary(ParVarMod)
```

```{r}
## Check with lme4 package
ParVarMod_lme4 <- glmer(Parasites ~ LizardType * SMI + Month + (1|PlotID),
                        family = "poisson",
                        data = morph_highDen
                      )
  
summary(ParVarMod_lme4)
```


<br/>
<br/>

#### Table S2
```{r TableS2: Models for subsampled data, echo=T, results='markup', warning=F}
## ParVarMod with alternate months - 1,3,5,7 to reduce replicate sampling
morph_highDen_altMonth1 <- morph_highDen %>% filter(Month %in% c(1,3,5,7)) 

ParVarMod_altMonth1 <- glmmadmb(Parasites ~ LizardType * SMI + Month,
                       random = ~1|PlotID , 
                       family = "poisson", 
                       zeroInflation = T,
                       data = morph_highDen_altMonth1)
summary(ParVarMod_altMonth1)

## ParVarMod with alternate months - 2,4,6
morph_highDen_altMonth2 <- morph_highDen %>% filter(Month %in% c(2,4,6)) 

ParVarMod_altMonth2 <- glmmadmb(Parasites ~ LizardType * SMI + Month,
                       random = ~1|PlotID , 
                       family = "poisson", 
                       zeroInflation = T,
                       data = morph_highDen_altMonth2)
summary(ParVarMod_altMonth2)
```
<br/>

### 3. Relationship between ectoparasites and visual signals
<br/>
```{r load-visualData}
##### Visual Colour Analysis
visual <- read.csv("/Users/apal/Projects/2015-Cnemaspis/analysis/datasets/visual_2023-05.csv", header=TRUE)
visual$PlotID_short <- as.factor(visual$PlotID_short)
str(visual)
```

```{r fig2, fig.height=4, fig.width=6}
fig2A <- ggplot (visual, aes(x=Parasites, y=Brightness)) +
    geom_point(shape=1, col="gold3") +
    labs(x="Ectoparasite Level", y="Brightness") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

fig2B <- ggplot (visual, aes(x=Parasites, y=Chroma)) +
    geom_point(shape=1, col="gold3") +
    labs(x="Ectoparasite Level", y="Chroma") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

fig2C <- ggplot (visual, aes(x = Parasites, y = Hue)) +
    geom_point(shape=1, col="gold3") +
    labs(x="Ectoparasite Level", y="Hue") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

grid.arrange(fig2A, fig2B, fig2C, ncol=3)
```
Fig2. Ectoparasite level vs visual signals

```{r SMI-for-visual-data}
# Visual mean SVL
meanSVL_vis <- mean(visual$SVL)
# Linear regression of log transformed mass by SVL
lm_vis <- lm(log(SVL)~log(Mass), visual)
summary(lm_vis)
slope_vis <- -0.02054

visual$SMI <- visual$Mass*(meanSVL_vis/visual$SVL)^slope_vis
```
<br/>
<br/>

#### Table S3
```{r Table S3: Models for visual signals, echo=T, warning=F, results='markup'}
modBright <- glmer(Brightness ~ Parasites*SMI + (1|PlotID_short), data = visual)
summary(modBright)
Anova(modBright)

modChroma <- glmer(Chroma ~ Parasites*SMI + (1|PlotID_short), data = visual)
summary(modChroma)
Anova(modChroma)

modHue <- glmer(Hue ~ Parasites*SMI + (1|PlotID_short), data = visual)
summary(modHue)
Anova(modHue)
```
<br/>

### 3. Relationship between ectoparasites and chemical signals
<br/>
```{r load-chemData}
chemical <- read.csv("./analysis/datasets/chemical_continuous.csv")
str(chemical)
```
<br/>

```{r fig3, fig.height=4, fig.width=4}

##proportion of aromatic compounds wrt ectoparasite load
fig3A <- ggplot (chemical, aes(x=Parasites, y=X.Aromatic)) +
    geom_point(shape=1, col="gold3") +
    labs(x="Ectoparasite Level", y="Aromatic Compounds (%)") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

##Total no. of compounds wrt ectoparasite load
fig3B <- ggplot (chemical, aes(x=Parasites, y=X.compounds)) +
    geom_point(shape=1, col="gold3") +
    labs(x="Ectoparasite Level", y="Total no. of compounds") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

grid.arrange(fig3A, fig3B, nrow = 1)
```
Fig3. Ectoparasite Levels vs chemical signals

<br/>

```{r figS3, fig.height=4, fig.width=2}

##proportion of aromatic compounds wrt ectoparasite load
figS3 <- ggplot (chemical, aes(x=Parasites, y=X.Saturated)) +
    geom_point(shape=1, col="gold3") +
    labs(x="Ectoparasite Level", y="Saturated Compounds (%)") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

figS3
```
figS3. Ectoparasite Levels vs percentage of saturated chemicals

<br/>

```{r fig S4}
##proportion of aromatic compounds wrt ectoparasite load
figS4a <- ggplot(subset(chemical, Parasites < 30), aes(x=Parasites, y=X.Aromatic)) +
    geom_point(shape=1, col="gold3") +
    labs(x="Ectoparasite Level", y="Aromatic Compounds (%)") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

##proportion of saturated compounds wrt ectoparasite load
figS4b <- ggplot (subset(chemical, Parasites < 30), aes(x=Parasites, y=X.Saturated)) +
    labs(x="Ectoparasite Level", y="Saturated Compounds (%)") +
    geom_point(shape=1, col="gold3") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

##Total no. of compounds wrt ectoparasite load
figS4c <- ggplot (subset(chemical, Parasites < 30), aes(x=Parasites, y=X.compounds)) +
    geom_point(shape=1, col="gold3") +
    labs(x="Ectoparasite Level", y="Total no. of compounds") +
    geom_smooth(method='lm', fullrange = TRUE, se=TRUE, level=0.95, span=1, 
                color="gold3", fill="gold3") +
    my_geom_theme

grid.arrange(figS4a, figS4b, figS4c, nrow = 1)
```

FigS3. Ectoparasite Levels vs percentage of saturated chemicals (without outliers)


#### Table S5
```{r Table S5: Models for chemical signals, echo=T, warning=F, error=F, results='markup'}
# % aromatic compounds as a function of parasite level:
aroma <- glmer(X.Aromatic ~ Parasites + (1|SMI), data = chemical)
summary(aroma)
Anova(aroma)

# % saturated compounds as a function of parasite level:
sat <- glmer(X.Saturated ~ Parasites + (1|SMI), data = chemical)
summary(sat)
Anova(sat)

# number of compounds as a function of parasite level:
comp <- glmer(X.compounds ~ Parasites + (1|SMI), data = chemical)
summary(comp)
Anova(comp)
```

#### Table S6
```{r Table S6: Models for chemical signals w/o outliers, echo=T, warning=F, error=F, results='markup'}
# % aromatic compounds as a function of parasite level:
aroma <- glmer(X.Aromatic ~ Parasites + (1|SMI), data = subset(chemical, Parasites < 30))
summary(aroma)
Anova(aroma)

# % saturated compounds as a function of parasite level:
sat <- glmer(X.Saturated ~ Parasites + (1|SMI), data = subset(chemical, Parasites < 30))
summary(sat)
Anova(sat)
 
# number of compounds as a function of parasite level:
comp <- glmer(X.compounds ~ Parasites + (1|SMI), data = subset(chemical, Parasites < 30))
summary(comp)
Anova(comp)
```

<br/>

### 4. Behavioural Analysis with visual and chemical signals

<br/>
```{r}
behav <- read.csv("./analysis/datasets/BehavioralData_sequential.csv", header=TRUE)
behav$Signal <- as.factor(behav$Signal)
behav$Treatment <- as.factor (behav$Treatment)
behav$FemaleID <- as.factor(behav$FemaleID)

str(behav)
```
<br/>

#### Table S7 (Visual)
```{r}
#hist(behav$RelFlick)
visFlick <- glmmadmb(abs(RelFlick)~Treatment, random = ~1|FemaleID, family="poisson", data=subset(behav, Signal=="Visual"))
summary (visFlick)
summary (glht(visFlick, linfct=mcp(Treatment="Tukey")))
```

```{r}
visMov <- glmmadmb(abs(RelMov)~Treatment, random=~1|FemaleID, family="poisson", data=subset(behav, Signal=="Visual"))
summary (visMov)
summary (glht(visMov, linfct=mcp(Treatment="Tukey")))
```


<br/>

#### Table S7 (Chemical)
```{r}
chemFlick <- glmmadmb(abs(RelFlick)~Treatment, random=~1|FemaleID, family="poisson", data=subset(behav, Signal=="Chemical"))
summary (chemFlick)
summary (glht(chemFlick, linfct=mcp(Treatment="Tukey")))
```

```{r}
chemMov <- glmmadmb(abs(RelMov)~Treatment, random=~1|FemaleID, family="poisson", data=subset(behav, Signal=="Chemical"))
summary (chemMov)
summary (glht(chemMov, linfct=mcp(Treatment="Tukey")))
```


#### Figure 4
```{r fig4}
plot_visFlick <- ggplot (subset(behav, Signal == "Visual"), aes(x=Treatment, y=RelFlick)) +
  geom_boxplot (size = 0.5, fill = "turquoise") +  geom_jitter (width = 0.1, shape = 21) +
  ylim (-7,10) + labs (x="", y="Relative Tongue Flicks", title = "Visual Signals") +
  scale_x_discrete (labels=c("Control 1", "Control 2", "Test")) +
  geom_hline(yintercept=0, linetype=2) + my_geom_theme + theme(plot.title = element_text(hjust = 0.5))

plot_chemFlick <- ggplot (subset(behav, Signal == "Chemical"), aes(x=Treatment, y=RelFlick)) +
  geom_boxplot (size = 0.5, fill = "indianred") +
  geom_jitter (width = 0.1, shape = 21) +
  ylim (-7,10) + 
  labs (x="", y="Relative Tongue Flicks", title = "Chemical Signals") +
  scale_x_discrete (labels=c("Control 1", "Control 2", "Test")) +
  geom_hline(yintercept=0, linetype=2) +
  my_geom_theme + theme(plot.title = element_text(hjust = 0.5))

plot_visApp <- ggplot (subset(behav, Signal == "Visual"), aes(x=Treatment, y=RelApp)) +
  geom_boxplot (size =  0.5, fill = "turquoise") +
  geom_jitter (width = 0.1, shape = 21) +
  ylim (-7,10) + 
  labs (x="Treatments", y="Relative Approach") +
  scale_x_discrete (labels=c("Control 1", "Control 2", "Test")) +
  geom_hline(yintercept=0, linetype=2) +
  my_geom_theme

plot_chemMov <- ggplot (subset(behav, Signal == "Chemical"), aes(x=Treatment, y=RelMov)) +
  geom_boxplot (size = 0.5, fill = "indianred") +
  geom_jitter (width = 0.1, shape = 21) +
  ylim (-7,10) + 
  labs (x="Treatments", y="Relative Movement") +
  scale_x_discrete (labels=c("Control 1", "Control 2", "Test")) +
  geom_hline(yintercept=0, linetype=2) +
  my_geom_theme

grid.arrange(plot_visFlick, plot_chemFlick, plot_visApp, plot_chemMov)
```

<br/>

### Simultaneous Behavioural Analysis with only chemical signals

<br/>
```{r}
behav_sim <- read.csv("analysis/datasets/BehavioralData_simultaneous.csv", header = T)
str(behav_sim)
```

#### Figure 5
```{r fig5}
ggplot(behav_sim, aes(fill=choice, x=type, y=count)) + geom_bar(position = "stack", stat="identity") +
  coord_flip() +theme_light()
```

